Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-10 Chapter-1/1-Theory.tex
s

Critical challenges await  the next decade of experimental neutrino physics. % will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector. 
%%% Plug to LArTPC %%%%%%%%%%%%
Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of these fundamental questions.  This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.

The main goals of these research programs include:
\begin{itemize}
\item[-] Assessment of the existence of right-handed sterile neutrinos. 
\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
\item[-] Determination the status of CP symmetry in the lepton sector.
\end{itemize}

@@

\section{Motivations for Hadronic Cross Sections in Argon}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% ---------------------------------------- Hadronic Cross sections ---------------------------------------- %%%%%
%%%%% ----------------------------------------     Pion Cross Section     ---------------------------------------- %%%%%

@@

\subsubsection{$\pi^{-}$Ar Cross section in the Context of Neutrino Searches}
As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino. In particular,  assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor. Understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant. For example, in case of resonant scattering 
\begin{equation}
\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N', 
\end{equation}
the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which will bounce within the nuclear medium. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.

\begin{table}[b]
\centering
\begin{tabular}{|l|l|}
\hline
$\Delta$ (1232) 3/2$^+$                 & $N^*$ (1440) 1/2$^+$    \\ \hline
$\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ & $N^* \rightarrow N \pi$ \\
$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  &  $N^* \rightarrow N \pi\pi$\\ 
$\Delta^{+} \text{(uud)} \rightarrow p \pi^0$  &                         \\ 
$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$  &                         \\ 
$\Delta^{0} \text{(udd)} \rightarrow p \pi^-$  &                         \\ 
$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  &                         \\ \hline
\end{tabular}
\caption{Main decay modes of the lightest Delta resonance and Nucleon excited state.}
\label{tab:decaysdelta}
\end{table}

The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation). As one of the most popular event generators for LAr experiments, we take GENIE~\cite{Andreopoulos:2009rq} here as an example.
The nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}. This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus. For A $ >$ 20 like argon, the 2-parameter Woods-Saxon shell model for density function is used.  The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus. This module places the outgoing particles in the nucleus and propagates them.  In the INRANUKE model, hadrons cannot undergo more than one FSI per event. When possible, external hadron-nucleus scattering data are used to tune INTRANUKE. Since no data is available for Argon, an interpolation of data from heavier and lighter nuclei is used for  the pion-argon cross section.

>>>>

@@

\section{Motivations for Hadronic Cross Sections in Argon}
Critical challenges await  the next decade of experimental physics at the intensity frontier. % will entail an even deeper understanding of the neutrino mixing pattern, investigation of the neutrino mass origin, the determination of the number of neutrinos and their nature, and the measurement of CP violation in the lepton sector. 
%%% Plug to LArTPC %%%%%%%%%%%%
Following the recommendation of the latest Particle Physics Project Prioritization Panel  \cite{P5}, the US  is dedicating substantial resources to the development of a short- and long- baseline neutrino program to address many of open questions in neutrino physics today.  This program pivots on the Liquid Argon Time Projection Chamber (LArTPC) detector technology which will be described in \ref{ch:1}.

The main goals of these research programs include:
\begin{itemize}
\item[-] Assessment of the existence of right-handed sterile neutrinos. 
\item[-] Determination of the sign of $\Delta m^2_{13}$ (or $\Delta m^2_{23}$), i.e., the neutrino mass hierarchy.
\item[-] Determination of the octant, i.e.  whether $\theta_{23}$ is maximal.
\item[-] Determination the status of CP symmetry in the lepton sector.
\item[-] \textcolor{red}{something about GUTs}.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% ---------------------------------------- Hadronic Cross sections ---------------------------------------- %%%%%
%%%%% ----------------------------------------     Pion Cross Section     ---------------------------------------- %%%%%

@@

\subsubsection{$\pi^{-}$Ar Cross section in the Context of Neutrino Searches}
As outlined in \ref{ch:NuInt}, neutrino experiments use the products of neutrino interactions to identify the energy and flavor of the incoming neutrino.
Pions are a common product of neutrino interaction, especially in resonant scattering, DIS and coherent pion production. For neutrino experiments in argon, there are two main reasons why understanding pion hadronic interactions with argon is important: to model the behavior of the pion inside the nucleus struck by the neutrino and to model the behavior of the pion during its propagation inside the detector medium.

Assumptions on the nuclear models and on the interaction of hadrons in the nucleus performed at the level of the neutrino event generator bridge the measurement of the products of a neutrino interaction to the reconstruction of the neutrino energy and flavor. Thus, understanding pion hadronic interactions with the nucleus is particularly important to model correctly resonant, DIS and coherent pion production in neutrino interactions, where the presence of pions in the nucleus is abundant. For example, in case of resonant scattering 
\begin{equation}
\nu_l + N \rightarrow l + \Delta/N^* \rightarrow l+ \pi + N', 
\end{equation}
the $\Delta$ and $N^*$ and excited states will decay hadronically in matters of $\sim10^{-24}$~s inside the nucleus producing pions which will bounce within the nuclear medium. The decay modes for the lower mass $\Delta$ (1232) and $N^*$(1440)  are listed in table \ref{tab:decaysdelta}.
\begin{table}[b]
\centering
\begin{tabular}{|l|l|}
\hline
$\Delta$ (1232) 3/2$^+$                 & $N^*$ (1440) 1/2$^+$    \\ \hline
$\Delta^{++} \text{(uuu)} \rightarrow p \pi^+$ & $N^* \rightarrow N \pi$ \\
$\Delta^{+} \text{(uud)} \rightarrow n \pi^+$  &  $N^* \rightarrow N \pi\pi$\\ 
$\Delta^{+} \text{(uud)} \rightarrow p \pi^0$  &                         \\ 
$\Delta^{0} \text{(udd)} \rightarrow n \pi^0$  &                         \\ 
$\Delta^{0} \text{(udd)} \rightarrow p \pi^-$  &                         \\ 
$\Delta^{-} \text{(ddd)} \rightarrow n \pi^-$  &                         \\ \hline
\end{tabular}
\caption{Main decay modes of the lightest Delta resonance and Nucleon excited state.}
\label{tab:decaysdelta}
\end{table}
The key elements of a neutrino event generators for resonance and DIS events are the nuclear model and the hadron treatment (both production and transportation). As one of the most popular event generators for LAr experiments, we take GENIE~\cite{Andreopoulos:2009rq} here as an example.
The nuclear model used by GENIE for all processes is a relativistic Fermi gas (RFG) modified to incorporate short range nucleon-nucleon correlations \cite{PhysRevD.24.1400}. This means that the initial momentum and binding energy of the struck nucleon is determined by assuming nucleons inside the nucleus are quasi-free, acting independently in the mean field of the nucleus. For A $ >$ 20 like argon, the 2-parameter Woods-Saxon shell model for density function is used.  The GENIE module INTRANUKE \cite{Juszczak2006} is used to simulate the final-state interactions (FSI) which is the hadron re-interaction inside the nucleus. This module places the outgoing particles in the nucleus and propagates them.  In the INRANUKE model, hadrons cannot undergo more than one FSI per event. When possible, external hadron-nucleus scattering data are used to tune INTRANUKE. Since no data is available for Argon, GENIE uses an interpolation of data from heavier and lighter nuclei for the pion-argon cross section leading to big uncertainties in the INTRANUKE module.
